---
title: "Chapter 5"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

### Example 5-1(1) Linear Regression with Dummy Variable

#### Input data
Y = BLOOD PRESSURE (SBP)                   
X = AGE                                    
Z = 0 FOR MALE, 1 FOR FEMALE 
```{r}
z <- c(0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1)
y <- c(158, 156, 176, 144, 150, 174, 165, 153, 144, 170, 150, 135, 142, 144, 185, 184, 164, 149, 156, 158, 172, 148, 138, 124, 140, 114, 120, 130, 152, 138, 154, 128, 134, 144, 160, 140, 145, 158, 110, 116, 120, 125, 159, 172, 124, 130, 134, 139, 157, 132, 162, 154, 128, 124, 160, 175, 176, 168, 142, 138, 174, 180, 170, 169, 142, 162, 130, 136, 158)
x <- c(41, 47, 65, 50, 38, 55, 56, 40, 39, 67, 56, 45, 50, 63, 60, 68, 57, 47, 52, 65, 62, 35, 45, 42, 59, 17, 39, 29, 41, 43, 61, 19, 41, 33, 51, 33, 47, 67, 34, 20, 21, 25, 47, 68, 36, 22, 18, 23, 48, 26, 65, 56, 42, 19, 44, 69, 66, 57, 44, 21, 51, 70, 59, 61, 46, 64, 48, 36, 53)
xz <- x*z

All <- data.frame(z,y,x,xz)
All$z <- factor(All$z, levels = c("0","1"), labels = c("0","1"))
str(All)
```

#### Date description
```{r}
library(psych)
describeBy(All[,2:3],All$z)
```

#### Regression model with dummy variable
```{r}
# Full model
model <- lm(y~.,data = All)

anova(model)
summary(model)
```

#### Hypothesis tests
```{r}
# test intercept beta(z) = 0
library(car)
linearHypothesis(model,c("z1 = 0"),test = "F")

# test slope beta(xz) = 0
linearHypothesis(model,c("xz = 0"),test = "F")

# test conincidence beta(z,xz) = 0
linearHypothesis(model,c("z1 = 0","xz = 0"),test = "F")

```

#### Scatterplot
```{r}
scatterplot(x~y | z, regLine=TRUE, smooth=FALSE, boxplots=FALSE,  by.groups=TRUE, legend=list(coords="bottomright"), data=All)
```


### EXAMPLE 5-1(2) Linear Regression with Dummy Variable

#### Input data
Y = BLOOD PRESSURE (SBP)                                    
X = AGE                                                     
Z = 0 FOR MALE, 1 FOR FEMALE                               
G = 1 FOR MALE, 2 FOR FEMALE  

```{r}
z <- c(0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1)
y <- c(158, 156, 176, 144, 150, 174, 165, 153, 144, 170, 150, 135, 142, 144, 185, 184, 164, 149, 156, 158, 172, 148, 138, 124, 140, 114, 120, 130, 152, 138, 154, 128, 134, 144, 160, 140, 145, 158, 110, 116, 120, 125, 159, 172, 124, 130, 134, 139, 157, 132, 162, 154, 128, 124, 160, 175, 176, 168, 142, 138, 174, 180, 170, 169, 142, 162, 130, 136, 158)
x <- c(41, 47, 65, 50, 38, 55, 56, 40, 39, 67, 56, 45, 50, 63, 60, 68, 57, 47, 52, 65, 62, 35, 45, 42, 59, 17, 39, 29, 41, 43, 61, 19, 41, 33, 51, 33, 47, 67, 34, 20, 21, 25, 47, 68, 36, 22, 18, 23, 48, 26, 65, 56, 42, 19, 44, 69, 66, 57, 44, 21, 51, 70, 59, 61, 46, 64, 48, 36, 53)
xz <- x*z
g <- z + 1
xg <- x*g
All <- data.frame(g,z,y,x,xz,xg)

All$z <- factor(All$z, levels = c("0","1"), labels = c("0","1"))
All$g <- factor(All$g, levels = c("1","2"), labels = c("1","2"))

str(All)
```


#### seperate regression model
```{r}
# z = 0
model1.1 <- lm(y ~ x, data = All[All$z==0,])
anova(model1.1)
summary(model1.1)
# z = 1
model1.2 <- lm(y ~ x, data = All[All$z==1,])
anova(model1.2)
summary(model1.2)
```


#### Regression model by z
```{r}
model2 <- lm(y ~ x + z + xz, data=All)
anova(model2)
summary(model2)
```


#### Regression model by g
```{r}
model3 <- lm(y ~ x + g + xg, data = All)
anova(model3)
summary(model3)
```



### EXAMPLE 5-2 ANALYSIS OF  COVARIANCE

#### Input data
```{r}
rm(list=ls())
trt <- c(1, 1, 1, 1, 2, 2, 2, 2, 3, 3, 3, 3, 4, 4, 4, 4, 5, 5, 5, 5)
rep <- c(1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3, 4)
initial <- c(27.2, 32, 33, 26.8, 28.6, 26.8, 26.5, 26.8, 28.6, 22.4, 23.2, 24.4, 29.3, 21.8, 30.3, 24.3, 20.4, 19.6, 25.1, 18.1)
final <- c(32.6, 36.6, 37.7, 31, 33.8, 31.7, 30.7, 30.4, 35.2, 29.1, 28.9, 30.2, 35, 27, 36.4, 30.5, 24.6, 23.4, 30.3, 21.8)

All <- data.frame(trt,rep,initial,final)
All$trt <- factor(All$trt)
All$rep <- factor(All$rep)
str(All)
```

#### Sample means
```{r}
# overall average
c(paste0("Mean of Initial: ",mean(All$initial)),
     paste0("Mean of Final: ",mean(All$final)))

# mean group by trt
aggregate(All[,3:4],list(All$trt),mean)
```

#### ANOVA
```{r}
model1 <- aov(final~trt,data=All)

summary(model1)

library(car)
Anova(model1,type="III")
```

#### Least Significant Difference (LSD)
```{r}
# install.packages("agricolae")
library(agricolae)
out1 <- LSD.test(model1,"trt",p.adj = "none") #p-adj ="none" is t-student
print(out1)
plot(out1)
```


#### ANCOVA
```{r}
model2 <- aov(final ~ trt + initial,data=All)

anova(model2) # Type I
Anova(model2, type="II")
```

#### LSD
```{r}
out2 <- LSD.test(model2,"trt",p.adj = "none") #p-adj ="none" is t-student
print(out2)
plot(out2)
```

```{r}
# least squre means of adjusted y
# install.packages("lsmeans")
library(lsmeans)
lsmeans(model2, ~trt, data=All)
```


