---
title: "Chapter 1"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

### EXAMPLE 1 - 1 Descriptive Statistics 

```{r}
# input x:
x <- c(25, 35, 55, 15, 40, 25, 55, 35, 45, 5, 20)

#discriptive statistics:
library(pastecs)
options(scipen = 200)
res <- stat.desc(x,norm = TRUE)
round(res,4)
```

```{r}
#Test for location:
t.test(x)
```

```{r}
#Test for normality：
library(fBasics)
#Shapiro-Wilk's test for normality
shapiroTest(x)
#Kolmogorov-Smirnov normality test
ks.test(x, "pnorm", mean(x), sd(x)) 
#Cramer–von Mises normality test
cvmTest(x)
#Anderson–Darling normality test
adTest(x)
```

```{r}
#Quantiles
quantile(x,c(1,0.99,0.95,0.90,0.75,0.50,0.25,0.10,0.05,0.01,0))
```
```{r}
#Plots
#Stem leaf
stem(x,scale = 2) # "scale" controls the plot length.

#boxplot
boxplot(x,main="Boxplot for x")

#Normal probability plot
qqnorm(x)

#Histogram and density plot
plot(density(x),main = "Description Statistics")
hist(x, freq=F, add=T)
```

### EXAMPLE 1 - 2 T Test 

#### TWO INDEPENDENT SAMPLES
```{r}
rm(list=ls())
Water <- c("BOT",	"BOT",	"BOT",	"BOT",	"BOT",	"BOT",	"SUR",	"SUR",	"SUR",	"SUR",	"SUR",	"SUR")	
Metal <- c(0.43,	0.266,	0.567,	0.531,	0.707,	0.716,	0.415,	0.238,	0.39,	0.41,	0.605,	0.609)
All <- data.frame(Water, Metal)
str(All)
# Input two vectors
BOT <- c(0.43,0.266,0.567,0.531,0.707,0.716)
SUR <- c(0.415,0.238,0.39,0.41,0.605,0.609)

#descreptive statistics by two groups
library(doBy)
summaryBy(Metal~Water, data = All, FUN = stat.desc)

#Normality tests
shapiro.test(BOT)
shapiro.test(SUR) 

#Equality of variances
var.test(BOT,SUR)

#Two independent sample t-test
t.test(BOT,SUR,var.equal = TRUE)
```

The null hypothesis of normality and variances equality of both BOT and SUR are not rejected. P-value of two independent t-test is far mor than 0.05.

#### One-way anova 
```{r}
glm <- lm(Metal~Water, data = All)
library(car)
summary(glm)

# Type I test
anova(glm)  

# Type III test
Anova(glm,type = "III",icontrasts = c("contr.sum", "contr.poly"))
```

#### Levene's Test for Homogeneity of Variance
Use the Levene’s test to check the homogeneity of variances. The function leveneTest() [in car package] will be used.
```{r}
library(car)
#ANOVA of absolute deviation from group means
leveneTest(Metal~Water, data = All,center = mean)
```
The p-value is not less than the significance level of 0.05. This means that there is no evidence to suggest that the variance across groups is statistically significantly different. Therefore, we can assume the homogeneity of variances in the different treatment groups.

#### Paired T test
```{r}
# Paired two sample t-test
dif <- stat.desc(BOT-SUR)
round(dif,4)
t.test(BOT-SUR)
```
P-value of paired t-test equals 0.014 which is bigger than 0.05. Not reject the null hypothesis.

### EXAMPLE 1 - 3 ONE-WAY ANOVA 

#### Data preparation
```{r}
rm(list=ls())
# Data input
TEA <- c('A', 'A', 'A', 'A', 'A', 'A', 'A', 'B', 'B', 'B', 'B', 'B', 'C', 'C', 'C', 'C', 'C', 'C', 'D', 'D', 'D', 'D', 'D', 'D')
VB <- c(7.9, 6.2, 6.6, 8.6, 8.9, 10.1, 9.6, 5.7, 7.5, 9.8, 6.1, 8.4, 6.4, 7.1, 7.9, 4.5, 5, 4, 6.8, 7.5, 5, 5.3, 6.1, 7.4)
TEA <- as.factor(TEA)
All <- data.frame(TEA,VB)
str(All)
# statistical description
library(dplyr)
All %>%
  group_by(TEA ) %>%
  summarise(N = n(),sum = sum(VB), mean = mean(VB),Var = var(VB))
summarise

```
#### The GLM procedure and One-way anova
```{r}
glm <- lm(VB~TEA, data = All)
library(car)
summary(glm)

# Type I test
anova(glm)  

# Type III test
Anova(glm,type = "III",icontrasts = c("contr.sum", "contr.poly"))
```

#### Levene's Test for Homogeneity of Variance
Use the Levene’s test to check the homogeneity of variances. The function leveneTest() [in car package] will be used.
```{r}
#ANOVA of absolute deviation from group means
leveneTest(VB, group = TEA, data = All,center = mean)
```

#### LSD test
```{r}
#install.packages("agricolae")
library(agricolae)
out <- LSD.test(glm,"TEA",p.adj = "none")
out
```

### EXAMPLE 1 - 4 Correlation Analysis 

#### Data preparation
```{r}
rm(list=ls())
y <- c(65, 78, 52, 82, 92, 89, 73, 98, 56, 75)
x <- c(39, 43, 21, 64, 57, 47, 28, 75, 34, 52)
All <- data.frame(y,x)
```

#### Data visualization
```{r}
# Scatterplot with star points
plot(All$x,All$y,xlab = "x",ylab="y",pch=8)
```

#### Correlation Analayis
```{r}
# install.packages("Hmisc")
library(Hmisc)

cor <- rcorr(All$x,All$y, type = "spearman")
cor

```










